---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dbutils

<!-- badges: start -->
``` {r}
use_lifecycle_badge("experimental")
```
<!-- badges: end -->

`dbutils` is a package of utility functions for processing and modifying BC Stats' population databases via three main categories:  

* database **access** functions (getDBPath, dbCheck, dbInfo, dbRead, dbWrite)  
* **conversion** functions (conversionRead, dbConvert)  
* **raking** functions (dbRake, raking helper functions)  


## About
The BC Stats Population section has historically depended on APL; the `dbutils` package is a conversion of the APL functions to R functions.

**Database** types include: population, deaths, births. Databases must have 7 columns: Year, Type, TypeID, Age, Male, Female, and Total. Each Year of data must have the same number of occurrences, and Years must be continuous (no year can be missing).

There are several **naming conventions** used with such data, where files are named **_DDDRRTYY_**:  

* **DDD** is the demographic type, such as BIR = Birth, DEA = Death, POP = Population, etc.  
* **RR** is the region code, such as:
  + RD - Regional District, which is the same data as Census Division (CD)  
  + DR - Development region (DR)  
  + HA - Local Heath Area (LHA)  
  + HS - Health Service Delivera Area (HSDA)  
  + HY - Health Authority (HA)  
  + CH - Community Health Service Area (CHSA)  
  + CF - Ministry of Children and Family Development (MCFD)  
  + CA - MCFD Service Delivery Area (MCFD_SDA)  
  + CL - MCFD Local Service Area (MCFD_LSA)  
  + SD - School District (SD)  
  + PS - College Region, or Post-Secondary (CR)  
  + SR - Special Regions 
* **T** is the data type, such as A = Actual, E = Estimates, O = Projections, P = Population  
* **YY** is the year (or two-digit PEOPLE version number, indicating which PEOPLE run) you want data for  


For example, population data is saved as POPRREYY (population estimates), POPRRPYY (population projections), BIRRRYY (births) or DEARRYY (deaths), where RR is the shorthand for the region code, and YY is the last two digits of the year. (This was passed over from decades of BC Stats PEOPLE development limited to 2-digit codes.) 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bcgov/dbutils")
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(dbutils)
## basic example code
```




